Legion: Expressing Locality and Independence with Logical Regions

### 2012-05-01 doc/sc2012/soop.tex
soop

\subsubsection{Dependences}
Conceptually, detecting dependences between a newly registered task $t$ and a previous task 
$t'$ requires comparing the two sets of logical regions accessed.  For each logical region used by
$t'$ that may overlap with a logical region used by $t$, the privileges
and coherence modes are compared to determine whether a dependence exists.  If
both regions need only read privileges there is never a dependence, but if either
task needs write or reduction privileges, the coherence modes are compared:
{\small
\begin{tabular}{c|cccc}
             & Exclusive & Atomic   & Simultaneous & Relaxed \\
\midrule
Exclusive    & Dep & Dep & Dep & Dep \\ 
Atomic       & Dep & Same & Cont & Cont \\
Simultaneous & Dep & Cont & Same & None \\
Relaxed      & Dep & Cont & None & None \\
\end{tabular}
}

A {\tt Dep} entry indicates dependence while {\tt None}
indicates independence.  A {\tt Same} entry is a dependence unless the two tasks
use the same physical instance of the logical region.\footnote{For example, if two tasks wish to acess the region with atomic coherence, then if they are working on the same physical copy of the data atomicity can be guaranteed using, e.g., locking.  If they are working on different copies of the data then the only safe execution strategy for Legion is to serialize the two tasks.}
A {\tt Cont} entry
indicates a dependence unless a single writer is using the
atomic coherence mode.  When a dependence is detected, the new task
waits for the older task to complete before it begins execution.

@@

\subsubsection{Isolation}
The reader may wonder why we only compare sibling tasks for
dependences: can't arbitrary pairs of tasks have dependences?  The
answer is ``yes'', but the following lemma shows that
any non-sibling tasks are always detected first as a dependence between
siblings that are registered earlier in execution!
\begin{lemma}
\label{lem:isolation}
\rm
Let $t_1$ and $t_2$ be two distinct, non-sibling tasks other than the root task with parents $p_1$ and $p_2$, respectively.
If $t_1$ and $t_2$ have a dependence, then $p_1$ and $p_2$ have a dependence.
\end{lemma}
\begin{proof}
Recall from Section~\ref{sec:ex} that the parents' region arguments include the children's region arguments and have
only more permissions.
\end{proof}
Lemma~\ref{lem:isolation} justifies checking dependences only between sibling tasks, because any dependences between 
arbitrary tasks are detected first between children of the tasks' least common ancestor.  Legion
guarantees that all dependence analysis is performed for all children of a task $t$ on the instance of the SOOP where $t$
runs, which by Lemma~\ref{lem:isolation} is sufficient to detect all required dependences across all tasks.

>>>>

\subsubsection{Dependences}
Detecting dependences between a newly registered task $t$ and a previous task 
$t'$ requires comparing the two sets of logical regions accessed.  For each logical region used by
$t'$ that may overlap with a logical region used by $t$, the privileges
and coherence modes are compared to determine whether a dependence exists.  If
both regions need only read privileges there is never a dependence, but if either
task needs write or reduction privileges, the coherence modes are compared using
the table in Figure~\ref{fig:dependence}.

\begin{figure}
{\small
\begin{tabular}{c|cccc}
             & Exclusive & Atomic   & Simultaneous & Relaxed \\
\midrule
Exclusive    & Dep & Dep & Dep & Dep \\ 
Atomic       & Dep & Same & Cont & Cont \\
Simultaneous & Dep & Cont & Same & None \\
Relaxed      & Dep & Cont & None & None \\
\end{tabular}
}
\caption{Dependence table.}
\label{fig:dependence}
\end{figure}
A {\tt Dep} entry indicates dependence while {\tt None}
indicates independence.  A {\tt Same} entry is a dependence unless the two tasks
use the same physical instance of the logical region.\footnote{For example, if two tasks wish to acess the region with atomic coherence, then if they are working on the same physical copy of the data atomicity can be guaranteed using, e.g., locking.  If they are working on different copies of the data then the only safe execution strategy for Legion is to serialize the two tasks.}
A {\tt Cont} entry
indicates a dependence unless a single writer is using the
atomic coherence mode.  When a dependence is detected, the new task
waits for the older task to complete before it begins execution.

@@

\subsubsection{Isolation}

\begin{figure}
\centering
\begin{tikzpicture}
\node(top) at (3.5,3.5) { $p$ };

\node(s1) at (1.9,2.5) { $s1$ };
\node(s2) at (5.1,2.5) { $s2$ };

\node(m1) at (1.9,1.5) { $\ldots$ };
\node(m2) at (5.1,1.5) { $\ldots$ };
\node(t1) at (1.9,0.5) { $t_1$ };
\node(t2) at (5.1,0.5) { $t_2$ };

\draw
  (top.south) edge (s1.north)
  (s2.north) edge (top.south)
  (m1.north) edge (s1.south)
  (m2.north) edge (s2.south)
  (t1.north) edge (m1.south)
  (t2.north) edge (m2.south)
  ;

\draw[dashed]
  (t1.east) edge (t2.west)
  (s1.east) edge (s2.west)
  ;

\end{tikzpicture}
\caption{Dependence between tasks implies a dependence between siblings of the least common ancestor.}
\label{fig:independence}
\end{figure}

The reader may wonder why we only compare sibling tasks for
dependences: can't arbitrary pairs of tasks have dependences?  The
answer is ``yes'', but the following lemma shows that
any non-sibling tasks are always detected first as a dependence between
siblings that are registered earlier in execution.
\begin{lemma}
\label{lem:isolation}
\rm
Let $t_1$ and $t_2$ be two distinct, non-sibling tasks other than the root task, let $p$ be the least
common ancestor task of $t_1$ and $t_2$, and let $s_1$ and $s_2$ be the sibling children of $p$ 
that are ancestors of $t_1$ and $t_2$, respectively.
If $t_1$ and $t_2$ have a dependence, then $s_1$ and $s_2$ have a dependence.
\end{lemma}
\begin{proof}
Recall from Section~\ref{sec:ex} that a task's parent, takes region arguments that include the children's region arguments and have only more permissions.  The same is true of any ancestor of a task.
\end{proof}
Lemma~\ref{lem:isolation} is illustrated in Figure~\ref{fig:isolation}:
if arbitrary tasks $t_1$ and $t_2$ have a dependence, then, because tasks always work on a subset
of the data passed to their parents, two siblings $s_1$ and $s_2$ of the task's
least common ancestor also have a dependence.  Since $s_1$ and $s_2$ are issued earlier in the program
execution than $t_1$ and $t_2$, $s_1$ and $s_2$ are serialized by the SOOP and the dependence between
$t_1$ and $t_2$ is observed.   Note that Lemma~\ref{lem:isolation} also guarantees a dual independence property: if two sibling tasks $s_1$ and $s_2$ have no dependence, then there can never be a dependence between the descendants of $s_1$ and the descendants of $s_2$.
